---
title: "Repo_research_project1"
author: "Sathya Thiruvengadam"
date: "May 2, 2019"
output: html_document
---

## This the code for reproduciable research project 1

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading needed package from library
```{r libini}
library(ggplot2)
library(dplyr)
library(lubridate)
```

Q.1 Read *.csv file and store in the variable "dfile"

```{r rdfile}
dfile <- read.csv("./activity.csv")
```


Q.2 Calculate total number of steps per day

```{r numofsteps}
numofsteps <- tapply(dfile$steps, dfile$date, sum, na.rm=TRUE)
```

Plot a histogram

```{r Numofsteps}
qplot(numofsteps, xlab='Total steps per day', ylab='Frequency using binwith 500', binwidth=500)
```

To calculate the mean and median of total number of steps per day
```{r meanmedian}
summary(numofsteps)
mean_numofsteps <- mean(numofsteps)
median_numofsteps <- median(numofsteps)
```

Q.3 Time series plot visualize the total number of steps, every day for an interval
```{r Avgstepsperinterval}
dfile %>% group_by(interval) %>%
    summarise(meanstepsperinterval = mean(steps, na.rm = TRUE)) %>%
    ggplot(aes(x=interval, y=meanstepsperinterval)) + geom_line() + labs(title = "Average steps per interval", x="Interval", y = "Avg steps") + theme(plot.title = element_text(hjust=0.5), plot.background = element_rect("aquamarine1"))
```

To get the interval with maximum steps every day
```{r maxstepsinterval}
dfile %>% group_by(interval) %>%
    summarise(meanstepsperinterval = mean(steps, na.rm=TRUE)) %>%
    filter(meanstepsperinterval == max(meanstepsperinterval)) 
```

Total number of rows with missing value NA in the dataset
```{r detectNAs}
colSums(is.na(dfile))
```

Q.4 Fill all NA's with column mean and create a new data frame "dfileNoNas" with missing data filled in.
```{r NAfillmean}
dfile %>% mutate_all(~ifelse(is.na(.x),mean(.x, na.rm=TRUE), .x)) -> dfileNoNAs
```

Plot a histogram for total number of steps every data with the Nas filled data
``` {r NumofstepsNoNAs}
TotalstepsdfileNoNAs <- aggregate(steps ~ date, data = dfileNoNAs, sum)
ggplot(TotalstepsdfileNoNAs, aes(x=steps)) + geom_histogram(binwidth = 500) +
    xlab("Total number of steps") +
    ggtitle("Total steps per day") +
    theme(plot.title = element_text(hjust = 0.5), plot.background = element_rect("aquamarine1"))
```

Calculate mean and median of the edited dataset
```{r meanmed}
summary(TotalstepsdfileNoNAs$steps)
Totalstepsmean=mean(TotalstepsdfileNoNAs$steps)
Totalstepsmedian = median(TotalstepsdfileNoNAs$steps)
```

Difference in distribution before and after missing values got filled-
We can see the distribution is symmetric once the missing values are filled,
mean is equal to median, where it was negative skewed before with mean less than median.

Q.5 Creating new variable DaysOfWeek using the function weekdays()
```{r detdaysofweek}
dfileNoNAs$date <- as.Date(dfileNoNAs$date, origin=lubridate::origin)
weekdays1 <- c("Monday","Tuesday","Wednesday","Thursday","Firday")
dfileNoNAs$DaysOfWeek <- factor((weekdays(dfileNoNAs$date) %in% weekdays1), 
                        levels=c(FALSE, TRUE),labels=c('weekend',"weekday"))
```

Generate a plot, showing the average number of steps based on interval
```{r Totalstepsbydaysofweek}
averageddfileNoNAs <- aggregate(steps ~ interval + DaysOfWeek, data=dfileNoNAs, mean)
ggplot(averageddfileNoNAs, aes(interval, steps)) + 
    geom_line() +
    facet_wrap(~DaysOfWeek,ncol=1) +
    labs(title="Steps per interval based on weekdays", x="Interval", y="Average number of steps")+ theme(plot.title = element_text(hjust=0.5), plot.background = element_rect("aquamarine1"))
```